---
title: "FCC providers list"
date: last-modified
draft: false
format:
  gfm:
    variant: +yaml_metadata_block
prefer-html: true
---

```{r}
#| label: utility functions
source("R/table_with_options.R")
```

We have multiple sources from FCC to define a provider. 

- Two are coming from the Broadband Data Collection:

  * The first one from [FCC](https://us-fcc.app.box.com/v/bdcprovideridtable), accessed the 26-04-2024

  * The second is comming from our pipelines of NBM (June 2023 release)

- One is comming from the Emergency Broadband Benefit[^ebb] and can be downloaded [here](https://www.fcc.gov/sites/default/files/ebb.provider.list_.xlsx) (accessed the 26-04-2024)    

[^ebb]: The Emergency Broadband Benefit is the precursor of the Affordable Connectivity Program, source: [https://www.fcc.gov/broadbandbenefit](https://www.fcc.gov/broadbandbenefit)

::: {.callout-tip}
- [cori.data.fcc](https://github.com/ruralinnovation/fcc.utils) is a small R package that contains providers data and is used here

- Data is downlable using the download button
:::
 
# What does this data look like?

## Broadband Data Collection versions:

```{r}
#|label: read csv
library("cori.data.fcc")

isp <- cori.data.fcc::fcc_provider
```

`FCC provider` has 4 341 rows and 5 columns.

Those columns are: 

- `Provider.Name`: Same than Brand Name?

- `Affiliation`: Same number than `Provider.ID`

```{r} 
#| column: margin
#| label: Operation type

op_type <-as.data.frame(table(isp$operation_type, dnn = "Type"), responseName = "Nb." )

knitr::kable(op_type)
```

- `Operation.Type`: Only two options "ILEC" or "Non-ILEC"

- `FRN`: **F**CC **R**egistration **N**umber; "number of the entity that submited the data". It is supposed to be a string of 10 characters (with padding 0). Slighly more number than `Provider.Name` and seems to be the primary key.

- `Provider.ID`: An ID for `Affiliations`

```{r}
#|label: display ISP from BDC
table_with_options(isp)
```

# EDA / Analysis 

## How many unique values do we have per column:

```{r}
#| label: count me some unique
#| column: margin
#| tbl-cap: Count of unique values / columns
sum_table <-  apply(isp, 2, function(x)length(unique(x))) |>
                    as.data.frame()
names(sum_table) <- c("Count of unique values")

knitr::kable(sum_table)
```

We can confirm that:

- `FRN` here is unique for every row in this data set (our primary key)

- We have a bit less `Provider.Name` (4321 / 4341) than `FRN`

- We have the same number of `Affiliations` and `Provider.ID`

- The number of `Provider.ID`/`Affiliations` is smaller than FRN. 

A quick check indicate that all `Provider.ID` are 6 characters.
`FRN` is also always 10 characters.

:::{.aside}
Hence the one with 7 characters in FCC NBM is probably an error.
:::

### What are the Provider Name that are sharing multiple `FRN`:

We probably have cases where companies have same name ("Farmers Mutual Telephone company") but we have probably company that have multiple FRN ("Grand Mound Cooperative Telephone Association"). Granted the low numbers I think we are fair to assume it does not matter to much. 

```{r}
#| label: case where same name have different FRN
provider.name_by_FRN <- sapply(split(isp$frn, isp$provider_name), function(x) unique(x))

multiple.frn <- provider.name_by_FRN[lengths(provider.name_by_FRN)> 1]

provider.name_by_FRN.dat <- data.frame(provider_name = names(multiple.frn),
                                   FRN = sapply(multiple.frn, toString)
)

table_with_options(provider.name_by_FRN.dat)
```


### How FRN are split between `Affiliations`/`Provider.ID`?

 As expected most of of the relations FRN / provider are one to one (3135) while 387 have more than one FRN. 

```{r}
#| label: spliting number of affiliation by frn
#| column: margin
#| tbl-cap: Affiliation per FRN
get_me_FRN_affiliations <- function(isp) { 
  FRN_by_affiliations <- sapply(split(isp[["frn"]], isp[["affiliation"]]), 
                                function(x) length(unique(x)))
  FRN_by_affiliations.dat <- data.frame(Affiliations = names(FRN_by_affiliations), 
                                        count_frn = FRN_by_affiliations)
  return(FRN_by_affiliations.dat)
}

FRN_affiliations.dat <- get_me_FRN_affiliations(isp) 
 
cnt_affiliation <- as.data.frame(table(FRN_affiliations.dat$count_frn),
                                       responseName = "")

cnt_affiliation[["Num. of Affiliations / FRN"]] <- 
  ifelse(as.numeric(cnt_affiliation[[1]]) < 10,  cnt_affiliation[[1]], "10+")

cnt_affiliation <- aggregate(cnt_affiliation[[2]], 
                             list(cnt_affiliation[["Num. of Affiliations / FRN"]]), 
                             sum)

names(cnt_affiliation) <- c("Number of Affiliations / FRN", "Count")

knitr::kable(cnt_affiliation[c(1, 3:9, 2 ),], row.names = FALSE)
```


You can explore those 387 affiliations here: 

```{r}
#| label: 387 cases wth one affiliation and multiple FR
dat <- FRN_affiliations.dat[FRN_affiliations.dat[["count_frn"]] > 1 ,]

table_with_options(dat[order(dat[["count_frn"]], decreasing = TRUE), ])
```


# Data set with mail address and phone numbers

This data set is one that is coming from the ACP program.

```{r}
#| label: address and phone
isp_contact <- read.csv("data/bb-provider_list.csv")
table_with_options(isp_contact)
```
