---
title: "EDA on 0/0 BSL"
date: last-modified
draft: false
format:
  gfm:
    variant: +yaml_metadata_block
prefer-html: true
---

TODO: update the SQL query

```{r}
#| label: utility functions
source("R/table_with_options.R")
# very lazish function, col should be a string 
agg_count <- function(dat, col) {
	agg <- aggregate(cbind(count = dat$count),
		             list(name_col = dat[[col]]),
					  sum)
	colnames(agg) <- c(col, "count")
	return(agg)
}
```

The goals of this page is storing a quick EDA about broadband services locations with 0 MBps uploads and 0 MBps downloads. To be concise we are going to call them 0/0 speeds. 

We have counted every services that have been declared with 0/0 speeds and associated with their States, ISP and technology. To clarify that does not mean a location have 0/0 speeds only but that one "ISP x technology" is provided with this kind of service in this location.  

The data used to provide most of the analysis was done with this 2 SQL queries. They were saved and stored in `data/`

```{sql}
#| label: SQL query to get the data 
#| eval: false
SELECT 
	state_abbr,
	brand_name,
	technology,
	count(brand_name)
FROM
	staging.june23
WHERE
(max_advertised_download_speed = 0 AND
	max_advertised_upload_speed = 0) = true
GROUP BY brand_name, state_abbr, technology;

-- first get all 0/0 then get all the non 0/0

SELECT 
	state_abbr,
	brand_name,
	technology,
	count(brand_name)
FROM 
	staging.june23
WHERE
(max_advertised_download_speed = 0 AND
	max_advertised_upload_speed = 0) = false
GROUP BY brand_name, state_abbr, technology;

```

```{r}
#| label: Load data
zero_loc <- read.csv("data/zero_dl_up.csv")
not_zero <- read.csv("data/not_zero_dl.csv")
```

## Summary by technologies:

```{r}
#| label: 0/0 by tecnology

agg <- agg_count(zero_loc, "technology") 
agg_not <- agg_count(not_zero, "technology")

technology <- merge(agg, agg_not, by.x = "technology", 
					by.y = "technology", all.x = TRUE, all.y = TRUE) 
colnames(technology) <- c("technology",  "cnt_zero_dl", "cnt_non_zero")
technology$rate_zero <- round(technology$cnt_zero_dl / 
				(technology$cnt_zero_dl +  technology$cnt_non_zero), 4)

table_with_options(technology)
```

</br>

We do not mind too much `70` (Unlicensed Terrestrial Fixed Wireless) because we are filtering it out but we are keeping `71` (Licensed Terrestrial Fixed Wireless) , `72` (Licensed-by-Rule Terrestrial Fixed Wireless)and `10` (Copper Wire). 

To take that into account I will filter out Unlicensed Terrestrial Fixed Wireless for the rest of this document. I also filtered out 60 and 61 to be consistant with our pipelines. 

## Summary by ISP

```{r}
#| label: ISP with 0/0
filter_sat <- c(60, 61, 70)
zero_loc <- zero_loc[which(! zero_loc$technology %in% filter_sat), ]
not_zero <- not_zero[which(! not_zero$technology %in% filter_sat), ]

agg <- agg_count(zero_loc, "brand_name")
agg_not <- agg_count(not_zero, "brand_name")

rate_zero <- merge(agg, agg_not, 
				   by.x = "brand_name", by.y = "brand_name"
				   , all.x = TRUE) 

colnames(rate_zero) <- c("brand_name", "cnt_zero_dl", "cnt_non_zero")
rate_zero$rate_zero <- round(rate_zero$cnt_zero_dl /
				 (rate_zero$cnt_zero_dl +  rate_zero$cnt_non_zero),
							4)

table_with_options(rate_zero[
					order(rate_zero$cnt_zero_dl, decreasing = TRUE),])
```

::: {.column-margin}
**402** ISPs are declaring services with 0/0 MBips (We have 2902 ISPs registered in FCC NBM)
:::

## Sumamry by States

```{r}
#| label: States
#| fig-column: margin 
st_agg_zero <- agg_count(zero_loc, "state_abbr")
st_agg_not <- agg_count(not_zero, "state_abbr")

st_agg <- merge(st_agg_zero, st_agg_not, 
				by.x = "state_abbr", by.y = "state_abbr",
				all.x = TRUE, all.y = TRUE)
colnames(st_agg) <- c("ST", "cnt_zero_dl", "cnt_non_zero")
st_agg$rate_zero <- round(st_agg$cnt_zero_dl /
				 	(st_agg$cnt_zero_dl +  st_agg$cnt_non_zero),
							4) 
library(ggplot2)

ggplot(st_agg[!is.na(st_agg$rate_zero),], aes(rate_zero)) +
  geom_boxplot(orientation = "y",
  fill='#A4A4A4', color="black") +
  coord_flip() +
  theme_bw()

table_with_options(st_agg[order(st_agg$rate_zero, decreasing = TRUE), ])

```


</br>
One point of concern is that services with 0/0 speeds could be generated for various reasons. One could be that some technology offer very low downloads/uploads and that is rounding to `0` an other could be that the location is not actually deserved but the ISP think it can do it.